Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-03 paper/paper.tex
minor changes

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of
messages. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri\ and
Section~\ref{sec:active-recon} shows a concrete example where \tname{}
uses this strategy. Although we built the option-field mapping
manually, from the \suri\ documentation, this is a one-time activity.
Unfortunately, for http attacks, the relevant data for the options is
likely in the message payload; not the fields. By analyzing our data
set of rules, we found that this occurs frequently in http attacks,
which are prevalent. For example, \percHttp\ of the \numrulessuri{}
rules from the ruleset we analyzed~\cite{emerging-threats-open}
prescribe defenses against http attacks. In addition,
\percRulesWithContent{} of these rules contain at least one content
option.

@@

Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
payload.  Figure~\ref{fig:distribution-contents} shows the
distribution of number of content options per rule for the data set
mentioned above. We limited the number of content options per rule to
8 in the histogram for space. \MyComment{Rules without
  \CodeIn{content} options are an exception and half of the rules
  contain at least two \CodeIn{content} options.}  To sum up, handling
content options for http is very important.

@@

\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.
\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ offiline by providing a \emph{pcap} file as input. \suri\ is invoked
multiple times for creating rules.

@@

  \renewcommand{\arraystretch}{0.35}

@@

%\renewcommand{\arraystretch}{0.7}
%\setlength{\tabcolsep}{2pt}
\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{2pt}
  \renewcommand{\arraystretch}{0.5}
  \caption{\label{table:results}Results.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
    \multicolumn{1}{c}{\multirow{2}{*}{\# Vars.}} &
    \#~Option Cand. &
    \#~Plausible Rules &    
    \multicolumn{1}{c}{Rank} \\

@@

\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{8pt}
  %\setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
      UDP Teardrop & 3 & 5$\uparrow$ & 2 \cellcolor{lightgray}$\downarrow$ & 3 & 3 \\
    ICMP Info: Ping & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    Dst/Host Unreachable & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    Dst/Net Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    Dst/Port Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    ICMP Info: Echo Reply & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    ICMP Info: Code 13 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    ICMP Info: Code 10 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2,665$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule in the ranking for an increasing number
\begin{wrapfigure}[11]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[10]{r}{0.28\textwidth}  
  \vspace{-2.5ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-2ex}
  \label{fig:impact-negative}
\end{wrapfigure}

>>>>

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of
messages. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri\ while
Section~\ref{sec:active-recon} shows a concrete example where \tname{}
uses this strategy to create the seed rule. Important to note that,
although we built the option-field mapping manually (from the
\suri\ documentation), this is a one-time activity.  Unfortunately,
this strategy is insufficient for http attacks, which are
prevalent. For example, \percHttp\ of the \numrulessuri{} rules from
the ruleset we analyzed~\cite{emerging-threats-open} prescribe
defenses against http attacks. By analyzing our data set of rules, we
found that the relevant data for the options appears in the message
payload, not the fields. For example, \percRulesWithContent{} of the
rules for http attacks contain at least one content option.

@@

Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of an
http rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.
Figure~\ref{fig:distribution-contents} shows the distribution of
number of content options per rule for the data set mentioned
above. We limited the number of content options per rule to 7 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  To sum up, content options appear
frequently in http rules and rules may contain several of these
options. Handling content options is therefore very important.

@@

\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.
\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
we calls \suri\ directly, through a public API, by providing a
\emph{pcap} files on input. \suri\ is invoked multiple times for
creating rules.

@@

  \renewcommand{\arraystretch}{0.4}

@@

%\renewcommand{\arraystretch}{0.7}
%\setlength{\tabcolsep}{2pt}
\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{2pt}
  \renewcommand{\arraystretch}{0.7}
  \caption{\label{table:results}Results.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
    \multicolumn{1}{c}{\multirow{2}{*}{\# Vars.}} &
    \#~Option Cand. &
    \#~Plausible Rules &    
    \multicolumn{1}{c}{Rank} \\

@@

\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{8pt}
  %\setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{0.5}
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
      UDP Teardrop & 3 & 5$\uparrow$ & 2 \cellcolor{lightgray}$\downarrow$ & 3 & 3 \\
    ICMP Info: Ping & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    Dst/Host Unreachable & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    Dst/Net Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    Dst/Port Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    ICMP Info: Echo Reply & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    ICMP Info: Code 13 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    ICMP Info: Code 10 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2,665$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule in the ranking for an increasing number
\begin{wrapfigure}[9]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-4ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[8]{r}{0.28\textwidth}  
  \vspace{-2.5ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-2ex}
  \label{fig:impact-negative}
\end{wrapfigure}
